---
title: "**Statistical Analyses**"
author: "Medha Tulsyani and Mukta Patwari"
output: 
  html_document: 
    code_folding: hide 
---

This page includes the statistical analyses performed to infer the relationship between AED density and traffic volume. The analyses performed were:

(1) Linear regression for mean traffic across boroughs,
(2) Linear regression for mean AEDs across boroughs,
(3) Correlation between AED density and mean traffic volume.

Additionally, a derived metric was created for the Traffic Burden Ratio to highlight the boroughs with the highest traffic congestion and lowest AED coverage.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(janitor)
library(plotly)
```

```{r import, include=FALSE}
## Importing the clean datasets

aed_inventory_df =
  read_csv("clean_data/aed_clean.csv")

leading_causes_clean =
  read_csv("clean_data/leading_causes_clean.csv")

traffic_volume_df =
  read_csv("clean_data/traffic_clean.csv")
```


### Linear model for mean traffic volume across boroughs, set Manhattan as reference
```{r}
traffic_volume_df$boro = as.factor(traffic_volume_df$boro)
traffic_volume_df$boro = relevel(traffic_volume_df$boro, ref = "Manhattan")
traffic_anova_model = lm(vol ~ boro, data = traffic_volume_df)
summary(traffic_anova_model)
```

On average, every borough has less traffic volume when compared to that of Manhattan.
On average, Manhattan has 147.6 cars worth of traffic volume at a given intersection.
The mean traffic volume in The Bronx is 61.1 cars less than that of Manhattan at a given intersection.
The mean traffic volume in Brooklyn is 24.9 cars less than that of Manhattan at a given intersection.
The mean traffic volume in Queens is 29.8 cars less than that of Manhattan at a given intersection.
The mean traffic volume in Staten Island is 38.6 cars less than that of Manhattan at a given intersection.

### Linear model for mean AED across boroughs, set Manhattan as reference
```{r}
aed_inventory_df$borough = as.factor(aed_inventory_df$borough)
aed_inventory_df$borough = relevel(aed_inventory_df$borough, ref = "Manhattan")
aed_anova_model = lm(aed_num_aeds ~ borough, data = aed_inventory_df)
summary(aed_anova_model)
```

On average, every borough has less AED density when compared to that of Manhattan.
On average, the AED density in Manhattan is 1.795 AED access points per location.
The mean AED density in The Bronx is .75 less AED access points per location compared to that of Manhattan.
The mean AED density in Brooklyn is .69 less AED access points per location compared to that of Manhattan.
The mean AED density in Queens is .32 less AED access points per location compared to that of Manhattan.
The mean AED density in Staten Island is .57 less AED access points per location compared to that of Manhattan.

### Pearson correlation between AED density and mean traffic volume

```{r}
mean_traffic =
  traffic_volume_df %>% 
  group_by(boro) %>% 
  summarize(mean_traffic = mean(vol, na.rm = TRUE))

aed_density = aed_inventory_df %>%
  group_by(borough) %>%
  summarize(total_aeds = sum(aed_num_aeds, na.rm = TRUE))

aed_traffic = mean_traffic %>% 
  full_join(aed_density, by = c("boro" = "borough"))

cor_result =
  cor.test(aed_traffic$total_aeds, aed_traffic$mean_traffic)

print(cor_result)
```

The resulting Pearson correlation coefficient is 0.826. This indicates a strong positive correlation between AED density and traffic volume, but the p-value is greater than 0.05, indicating it is non-significant. 

### Derived metric
```{r}
traffic_summary = traffic_volume_df %>%
  group_by(boro) %>%
  summarize(total_vehicles = sum(vol, na.rm = TRUE))

tbr_df = traffic_summary %>%
  full_join(aed_density, by = c("boro" = "borough"))

tbr_df = tbr_df %>%
  mutate(TBR = total_vehicles / total_aeds) %>%
  arrange(desc(TBR))

tbr_df
```

The Traffic Burden Ratio seems to be highest in Brooklyn and lowest in Manhattan. Although traffic is high in Manhattan, there is increased AED availability which mitigates the potential for a high TBR.