---
title: "Inferential Analyses"
author: "Medha Tulsyani and Mukta Patwari"
output: html_document
date: "2025-11-29"
---

Load Libraries
```{r}
library(tidyverse)
```



Import Data
```{r}
traffic_volume_df = read.csv("clean_data/traffic_clean.csv")

aed_inventory_df = read.csv("clean_data/aed_clean.csv")
```


ANOVA for mean traffic volume across boroughs
```{r}
traffic_anova_model = aov(vol ~ boro, data = traffic_volume_df)
summary(traffic_anova_model)
```

At least one borough's mean traffic volume differs from the others. This is significant and there is extremely strong evidence since the F value is 1,094.

ANOVA for mean AED across boroughs
```{r}
aed_anova_model = aov(aed_num_aeds ~ borough, data = aed_inventory_df)
summary(aed_anova_model)
```

At least one borough's mean AED differs from the others. This is significant with an F value of 24.09.

Linear model for mean traffic volume across boroughs, set Manhattan as reference
```{r}
traffic_volume_df$boro = as.factor(traffic_volume_df$boro)
traffic_volume_df$boro = relevel(traffic_volume_df$boro, ref = "Manhattan")
traffic_anova_model = lm(vol ~ boro, data = traffic_volume_df)
summary(traffic_anova_model)
```

On average, every borough has less traffic volume when compared to that of Manhattan.

Linear model for mean AED across boroughs, set Manhattan as reference
```{r}
aed_inventory_df$borough = as.factor(aed_inventory_df$borough)
aed_inventory_df$borough = relevel(aed_inventory_df$borough, ref = "Manhattan")
aed_anova_model = lm(aed_num_aeds ~ borough, data = aed_inventory_df)
summary(aed_anova_model)
```

On average, every borough has less AED density when compared to that of Manhattan.